---
title: "Observer data"
author: "Alberto Rovellini"
date: "9/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rbgm)
library(sf)
library(viridis)
library(maps)
library(mapdata)
library(data.table)
library(RColorBrewer)
```

```{r}
select <- dplyr::select
```

This document explores observer data from [AKFIN](https://akfinbi.psmfc.org/analytics/saw.dll?Dashboard). The data is called [NORPAC Catch Report](https://akfinbi.psmfc.org/analytics/saw.dll?Dashboard&PortalPath=%2fshared%2fStock%20Assessment%2f_portal%2fStock%20Assessment&Page=NORPAC%20Catch%20Report&Done=Dashboard%26PortalPath%3d%252fshared%252fStock%2520Assessment%252f_portal%252fStock%2520Assessment%26Page%3dObserver%2520and%2520EM%2520Data%26ViewState%3dt8rvrc7c2svggcggmbmq9ce062). Here are the Report Criteria:

* Year: 1989-2019 (in smaller chunks)
* FMP Area: GOA
* FMP Subarea: --Select Value--
* NMFS Area: --Select Value--
* Gear Code: --Select Value--
* Gear Description: --Select Value--
* Performance: --Select Value--
* Performance Description: --Select Value--
* Species Code: --Select Value--
* Species Name: --Select Value--

There seem to be some issues with AKFIN not picking the correct years that we select. Let's use distinct().

Note: as of September 2022, NORPAC Observer Catch Report data contains haul and trip target fields.

Read data.
```{r}
all_files <- list.files('../data/AKFIN/Observer_2022/',full.names = TRUE)
obs_list <- lapply(all_files, read.csv, skip=6)

obs <- rbindlist(obs_list)

# distinct rows
obs <- obs %>% distinct()

glimpse(obs)
```

The minimal unit in this data is the haul.

Can we make sense of what "Extrapolated weight" means here? Does it mean biomass in the haul based on the sample? Sum these for each species within a haul, and compare with the official catch.

```{r}
obs %>% group_by(Year,Haul.Join,Official.Total.Catch..mt.) %>% 
  summarise(Tot.weight = sum(Extrapolated.Weight..kg.,na.rm=T)) %>% 
  ungroup() %>% 
  #filter(Tot.weight!=0) %>% 
  #filter(Year==1989) %>%
  ggplot(aes(x=Official.Total.Catch..mt.,y=Tot.weight,color=factor(Year)))+
  theme_minimal()+
  geom_point()
```

It appears that the extrapolated weight is the weight of of a species in a haul as estimated from the subsample taken by the observer. That is, an approximation of the catch in the haul. The 0's must mean no estimate was taken but a catch was still reported.

# Species, gears, targets

Exploring what gears, targets, and species are explored in this data. This is useful to guide fleet structure determination.

## Gears
```{r}
obs %>% pull(Gear.Description) %>% unique()
```
Of the gear definitions, 'NON PELAGIC" is interesting. Does it mean bottom trawl?

## Targets

These are identified as haul and trip target. It is not a simple definition - some boats change haul target on the same trip, some boats process at sea so the concept of "trip" is not so defined in space, and boats will have different targets on different trips.

Here are haul targets:
```{r}
obs %>% pull(Haul_Target_Name) %>% unique() %>% sort()
```

And here are the trip targets:
```{r}
obs %>% pull(Trip_Target_Name) %>% unique() %>% sort()

```

How variable are haul targets within a trip target?
```{r}
# make plots for methods
colourCount <- length(obs %>% pull(Haul_Target_Name) %>% unique())
getPalette <- colorRampPalette(brewer.pal(12, "Paired"))

p <- obs %>%
  group_by(Trip_Target_Name, Haul_Target_Name) %>%
  tally() %>%
  drop_na() %>%
  group_by(Trip_Target_Name) %>%
  mutate(Prop = n/sum(n)) %>%
  ungroup() %>%
  ggplot()+
  geom_bar(aes(x = Trip_Target_Name, y = Prop, fill = Haul_Target_Name), stat = 'identity')+
  scale_fill_manual(values = getPalette(colourCount))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 60,hjust = 1,size = 12))
p

ggsave('../output/graphics/fleets/trip_vs_haul_targets.png', p, width = 8, height = 6)


```

Hauls can have targets different from the trip.

What are the target-gear combinations in the data? 
```{r}
combo <- obs %>%
  filter(Trip_Target_Name != '') %>% # drop this for now - does ti mean that a target was not assigned?
  mutate(Target_gear_combo = paste(Trip_Target_Name, Gear.Description, sep = '_')) %>%
  group_by(Year, Target_gear_combo) %>%
  summarise(Catch_mt = sum(Official.Total.Catch..mt.)) %>%
  group_by(Target_gear_combo) %>%
  summarise(Mean_annual_catch_mt = mean(Catch_mt)) %>%
  ungroup() %>%
  arrange(desc(Mean_annual_catch_mt))

write.csv(combo, 'target_gear_combo.csv', row.names = F)

p1 <- combo %>%
  ggplot()+
  geom_bar(aes(x = reorder(Target_gear_combo, -Mean_annual_catch_mt), y = Mean_annual_catch_mt), stat = 'identity')+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 60,hjust = 1,size = 12))

p1
ggsave('../output/graphics/fleets/target_gear_combinations.png', p1, width = 14, height = 5)
  
```

Species are caught with different gears, but the most effective fisheries are the ones we know about (rockfish trawl, pollock pelagic trawl, PCod and sablefish longline, pcod trawl and pot, flatfish trawl, etc.). There are some weird cases (like Pollock - bottom_PELAGIC, which is it??).

Finally, what species do we have data for?
```{r}
all_species <- obs %>% select(Species.Name) %>% distinct() %>% arrange(Species.Name) %>% pull()
all_species

obs_species <- obs %>% filter(Species.Name == "SABLEFISH (BLACKCOD)")
```

It looks like each Haul.Join denotes one row only, so no need to aggregate further for the hauls.

# Spatial resolution

Look at this in space? Then sum it all up for a species and compare with catch data from AKFIN. This last step will need aggregation by area, which we have in this dataset so that should be possible.
```{r}
obs_species_thin <- obs_species %>% select(Year,NMFS.Area,Gear.Description,Official.Total.Catch..mt.,Lat.DD.End,Lon.DD.End,X..Retained,Species.Name,Sample.Number:Extrapolated.Weight..kg.)

glimpse(obs_species_thin)
```

Visualise in space.
```{r}
obs_species_thin_sf <- obs_species_thin %>% st_as_sf(coords = c(x = 'Lon.DD.End', y='Lat.DD.End'), crs=4326)

obs_bbox <- obs_species_thin_sf %>% st_bbox()

coast <- maps::map("worldHires", c("USA","Canada"), plot = FALSE, fill = TRUE)
coast_sf <- coast %>% st_as_sf() #%>% st_transform(crs = atlantis_crs)

# pick a year and view the data for that year
obs_species_thin_sf %>% #filter(Year==2015) %>%
  ggplot()+
  geom_sf(aes(color=log1p(Extrapolated.Weight..kg.)))+
  geom_sf(data = coast_sf)+
  coord_sf(xlim = c(obs_bbox$xmin,obs_bbox$xmax), ylim = c(obs_bbox$ymin,obs_bbox$ymax))+
  scale_color_viridis()+
  theme_minimal()+
  facet_wrap(~Year)+
  labs(title = paste('Hauls from Observer data for',obs_species_thin_sf$Species.Name[1],sep = ' '))
```

Is there a spatial footprint of fishing fleets?
```{r}
top_fleets <- combo %>% slice_head(n = 8) %>% pull(Target_gear_combo) 

for(i in 1:length(top_fleets)){
  this_fleet <- top_fleets[i]
  
  p2 <- obs %>%
    filter(Trip_Target_Name != '') %>% # drop this for now - does ti mean that a target was not assigned?
    mutate(Target_gear_combo = paste(Trip_Target_Name, Gear.Description, sep = '_')) %>% 
    filter(Target_gear_combo == this_fleet & Official.Total.Catch..mt. > 0) %>%
    st_as_sf(coords = c(x = 'Lon.DD.End', y='Lat.DD.End'), crs=4326) %>%
    ggplot()+
    geom_sf(aes(color=log1p(Official.Total.Catch..mt.)))+
    geom_sf(data = coast_sf)+
    coord_sf(xlim = c(obs_bbox$xmin,obs_bbox$xmax), ylim = c(obs_bbox$ymin,obs_bbox$ymax))+
    scale_color_viridis()+
    theme_bw()+
    facet_wrap(~Year, ncol = 4)+
    labs(title = this_fleet)
  
  ggsave(paste0('../output/graphics/fleets/',this_fleet, '.png'), p2, width = 10, height = 14)
  
}

```












# Comparison with catch data

Catch estimates from the AKRO/BLEND Catch Accounting system integrate observer data, tickets, logbook data, etc. Comparing the two is probably not meaningful.

Read in catch data, and subset to the same years.
```{r}
catch <- read.csv("../data/AKFIN/Catch/Groundfish Total Catch.csv", fileEncoding = 'UTF-8-BOM')

catch_species <- catch %>% filter(Species.Group.Name == 'Pollock')
catch_species <- catch_species %>% filter(Year %in% (obs_species_thin %>% select(Year) %>% distinct() %>% pull()))

# remove discards
catch_species <- catch_species %>% filter(Retained.Discarded=='Retained')
```

Add these up by year and area and see how they plot.
```{r}
obs_species_by_area <- obs_species_thin %>% group_by(Year,NMFS.Area) %>% summarise(Area.Catch.Obs = sum(Official.Total.Catch..mt.))

catch_species_by_area <- catch_species %>% group_by(Year,NMFS.Area) %>% summarise(Area.Catch.Catch = sum(Catch..mt.))

# join

catch_comp <- obs_species_by_area %>% left_join(catch_species_by_area, by=c('Year','NMFS.Area'))

catch_comp %>%
  ggplot()+
  geom_point(aes(x=Area.Catch.Catch,y=Area.Catch.Obs))+
  geom_abline(intercept = 0,slope=1)+
  theme_minimal()
```


The catch data are from /Catch Accounting System. From [the CAS webpage](https://www.fisheries.noaa.gov/inport/item/21988), we read: *"The CAS uses a combination of industry reports and onboard observer information to provide an estimate of total catch and bycatch. Industry reported data consists of catch (landing reports) and processed product amounts (production reports), and these reports are electronically recorded and submitted to NMFS via eLandings. Other sources of information come from the Alaska Commercial Fisheries Entry Commission (CFEC), which issues permits and vessel licenses, and Vessel Monitoring Systems (VMS), which collect the position, time at a position, and course and speed of fishing vessels."*. Similarly, anout [BLEND](https://www.fisheries.noaa.gov/inport/item/11812): *"The Blend system used a combination of industry reports and observer data. For shoreside processors, Weekly Production Reports (WPR) submitted by industry were considered the best source of data for retained groundfish landings. All fish delivered to shoreside processors were weighed on scales, and these weights were used to account for retained catch. Observer data from catcher vessels provided the best data on at-sea discards of groundfish by vessels delivering to shoreside processors. Discard rates from these observer data were applied to the shoreside groundfish landings to estimate total at-sea discards from both observed and unobserved catcher vessels. For observed catcher/processors and motherships, the WPR and the Observer Reports recorded estimates of total catch (retained catch plus discards). If both reports were available, the Blend System selected one of them for incorporation into the catch database. If the vessel was unobserved, only the WPR was available."*.

The way people go from observer data to catch data was described to me as a "black box", so not sure there is hope for us to cleanly convert between the two here. There seem to be a correlation between them, at least. We may try and run this functionally for all species and see of there is a correlation, and then simply apportion the catch from the "Catch By Fishery" data frame to the boxes (where possible) based on whatever little data we have from this observer database.
