---
title: "AKFIN catch by fishery"
author: "Alberto Rovellini"
date: "9/1/2021"
output: html_document
---

This document explores the data set [Groundfish Total Catch by Fishery](https://akfinbi.psmfc.org/analytics/saw.dll?Dashboard&PortalPath=%2fshared%2fStock%20Assessment%2f_portal%2fStock%20Assessment&Page=Groundfish%20Total%20Catch%20by%20Fishery&Done=Dashboard%26PortalPath%3d%252fshared%252fStock%2520Assessment%252f_portal%252fStock%2520Assessment%26Page%3dCatch%2520Data%26ViewState%3dg5ebu3r609alcptvc3dh209ftq) from AKFIN. See AKFIN_catch.Rmd for details on the data query from AKFIN.

Objectives are:

1. Understand what "Trip target group" means.
2. Get a sense of how many species we have. 
3. Visualise aggregated catch by speies by NMFS area.
4. Map species and targets to the Atlantis groups and repeat.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(sf)
library(viridis)
library(lubridate)
library(maps)
library(mapdata)
library(data.table)
```

Read all files.
```{r}
all_files <- list.files('AKFIN/Catch/by_fisheries/',full.names = TRUE)
file_list <- lapply(all_files, read.csv,fileEncoding='UTF-8-BOM')
catch_byf <- rbindlist(file_list)

glimpse(catch_byf)
```

What are the targets here?
```{r}
all_target_groups <- catch_byf %>% select(Trip.Target.Group) %>% distinct() %>% pull()
all_target_names <- catch_byf %>% select(Trip.Target.Name) %>% distinct() %>% pull()

all_target_groups
all_target_names
```
Target groups are limited to main groundfish species. What are the gears?
```{r}
all_gears <- catch_byf %>% select(Gear) %>% distinct() %>% pull()
all_gears
```

Only codes are offered here, and they seem to be a subset of the "catch" data from AKFIN. Let's have a look at what gear is reported for flatfish - it should be bottom trawl.
```{r}
catch_byf %>% filter(Trip.Target.Name=="Pollock - midwater") %>% select(Gear) %>% distinct() %>% pull()
```
Gear names in the oberver data seem to be less cryptic (e.g. there is a long name field). 

What are the species recorded in the catch data?
```{r}
all_species_groups <- catch_byf %>% select(Species.Group) %>% distinct() %>% pull()
all_species <- catch_byf %>% select(Species) %>% distinct() %>% pull()

all_species_groups
all_species
```
The bycatch information from this data set seem to focus on the groups of commercial and conservation interest. For example, there are no benthic invertebrates nor seabirds, which are instead included in the observer data to some extent. 

# Map to NMFS areas

Let's calculate monthly catch based on NMFS statistical areas. It seems that data from here matches data from the annual aggregated catch (see AKFIN_catch.Rmd for detials on that).
```{r}
catch_byf <- catch_byf %>% mutate(Month = month(Week.Ending.Date))

catch_by_month <- catch_byf %>% group_by(Year,Month,Species.Group,Species,FMP.Area,FMP.Subarea,NMFS.Area) %>% summarise(Catch = sum(Catch..mt.,na.rm=FALSE)) %>% ungroup() %>% mutate(Month.Year = make_date(year = Year, month = Month))
```

Plot over time, by year for ease of visualisation.
```{r,fig.width=10,fig.height=20}
catch_by_month %>% 
  group_by(Year,Species.Group,Species,NMFS.Area) %>%
  summarise(Catch = sum(Catch, na.rm=TRUE)) %>%
  ungroup() %>%
  ggplot(aes(x=Year,y=Catch,fill=Species.Group))+
  geom_bar(stat='identity',position='stack')+
  theme_minimal()+
  guides(fill=guide_legend(ncol=1))+
  facet_wrap(~NMFS.Area, ncol=1)
```

Read in the shapefile for the NMFS areas.
```{r}
nmfs_sf <- st_read("gf95_nmfs/gf95_nmfs.shp")
nmfs_sf <- nmfs_sf %>% filter(NMFS_AREA %in% (catch_byf %>% select(NMFS.Area) %>% distinct() %>% pull()))

#limits for plotting
nmfs_bbox <- nmfs_sf %>% st_bbox()

# add some coastline
coast <- maps::map(database = "worldHires", regions = c("Canada","US"), plot=FALSE, fill=TRUE)
coast_sf <- coast %>% st_as_sf(crs = 4326) %>% st_transform(crs=st_crs(nmfs_sf)$proj4string)
```

Plot for a species, by area, over time (years).
```{r,fig.width=12,fig.height=20}
this_species <- 'Pacific Cod'

nmfs_sf %>% 
  left_join((catch_by_month %>% 
               filter(Species.Group==this_species) %>%
               group_by(Year,Species.Group,Species,NMFS.Area) %>%
               summarise(Catch = sum(Catch, na.rm=TRUE)) %>%
               ungroup()),
            by=c('NMFS_AREA'='NMFS.Area')) %>%
  ggplot()+
  geom_sf(aes(fill=Catch))+
  scale_fill_viridis()+
  geom_sf(data=coast_sf,fill=NA)+
  coord_sf(xlim=c(nmfs_bbox$xmin,nmfs_bbox$xmax),ylim=c(nmfs_bbox$ymin,nmfs_bbox$ymax))+
  theme_minimal()+
  facet_wrap(~Year)+
  labs(title = paste("Catch for",this_species,"by NMFS area"), fill="Catch (mt)")
  
```

This will need to be mapped to the Atlantis groups.