---
title: "AKFIN catch data"
author: "Alberto Rovellini"
date: "8/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(sf)
library(viridis)
library(lubridate)
```

This document explores the catch data from AKFIN. It also explores the differences between Catch and Catch by Fishery data sets. 

Both data sets come from the [Catch Data](https://akfinbi.psmfc.org/analytics/saw.dll?Dashboard&PortalPath=%2fshared%2fStock%20Assessment%2f_portal%2fStock%20Assessment&Page=Catch%20Data&Done=Dashboard%26PortalPath%3d%252fshared%252fStock%2520Assessment%252f_portal%252fStock%2520Assessment%26Page%3dGroundfish%2520Total%2520Catch%2520by%2520Fishery%26ViewState%3d4vmps7tab03fi787p32gt928p6) tab on AKFIN. Search details for each data set:

1. [Groundfish Total Catch](https://akfinbi.psmfc.org/analytics/saw.dll?Dashboard&PortalPath=%2fshared%2fStock%20Assessment%2f_portal%2fStock%20Assessment&Page=Groundfish%20Total%20Catch&Done=Dashboard%26PortalPath%3d%252fshared%252fStock%2520Assessment%252f_portal%252fStock%2520Assessment%26Page%3dCatch%2520Data%26ViewState%3dg989ikgqgnk97vlvv3ut9e0s76). Report criteria:

* Year: 1991-2021
* FMP Area: GOA
* FMP Subarea: --Select Value--
* Gear: --Select Value--
* Species Group: --Select Value--
* Choose a Report: "Detail with Processor/Vessel Characteristics"

2. [Groundfish Total Catch by Fishery](https://akfinbi.psmfc.org/analytics/saw.dll?Dashboard&PortalPath=%2fshared%2fStock%20Assessment%2f_portal%2fStock%20Assessment&Page=Groundfish%20Total%20Catch%20by%20Fishery&Done=Dashboard%26PortalPath%3d%252fshared%252fStock%2520Assessment%252f_portal%252fStock%2520Assessment%26Page%3dCatch%2520Data%26ViewState%3didnmhgqo8th0dfbdhbf798h6mi). Report criteria:

* Year: 1991-2021 (but in 6-7 year chunks, or else the data export chokes)
* Target Group: --Select Value--
* Target Name: --Select Value--
* FMP Area: GOA
* FMP Subarea: --Select Value--
* NMFS Area: --Select Value--
* FMP Gear: --Select Value--
* Species Group: --Select Value--
* Species: --Select Value--
* Choose a Report: "Detail with Processor/Vessel Characteristics"

All exports from AKFIN are done as .csv.

```{r}
catch <- read.csv("AKFIN/Catch/Groundfish Total Catch.csv", fileEncoding = 'UTF-8-BOM')

glimpse(catch)
```

What gears?
```{r}
all_gears <- catch %>% select(Gear) %>% distinct() %>% pull()
all_gears
```

What species do we have this for?
```{r}
all_species <- catch %>% select(Species.Group.Name) %>% distinct() %>% arrange(Species.Group.Name) %>% pull() 
all_species
```

Now read in the Catch By Fishery data. This takes substantially longer than the Catch data, perhaps because I chose the version with vessel detail for download.
```{r}
catch_byf <- read.csv("AKFIN/Catch/by_fisheries/Groundfish Total Catch by Fishery 2008-2014.csv", fileEncoding = 'UTF-8-BOM')

glimpse(catch_byf)
```

This is detailed information on the catch of a species or species group (e.g. sharks) by vessel. Notably, this data seems to have information at week level. It may need to be downloaded and processed year-by-year because of the size of it. It will still not map to anything smaller than the subareas.

Try for one species, see how this compares to the By Fishery data frame.
```{r}
species <- 'Atka Mackerel'

catch_species <- catch %>% filter(Species.Group.Name==species)
catch_species_byf <- catch_byf %>% filter(Species.Group==species)

# subset to years that we have catch by species for - this will change when we get all the data we need
catch_species <- catch_species %>% filter(Year %in% (catch_byf %>% select(Year) %>% distinct() %>% pull()))
```

```{r}
catch_species_summary <- catch_species %>% group_by(Year,NMFS.Area) %>% summarise(catch_sp = sum(Catch..mt.))
catch_species_byf_summary <- catch_species_byf %>% group_by(Year,NMFS.Area) %>% summarise(catch_sp_byf = sum(Catch..mt.))

catch_join <- catch_species_summary %>% left_join(catch_species_byf_summary,by=c('Year','NMFS.Area'))
```

```{r}
catch_join %>% ggplot()+
  geom_point(aes(x=catch_sp,y =catch_sp_byf))+
  geom_abline(intercept=0,slope=1)+
  theme_minimal()
```

The By Species data seems to be more fine-grained / less aggregated. Also it is wekly data, which is great.

Let's look more into the By Fishery data.
```{r}
glimpse(catch_byf)

# how many months?
levels(factor(month(ymd(catch_byf %>% select(Week.Ending.Date) %>% distinct() %>% pull())))) # data from all months
```

