---
title: "Fish Tickets"
author: "Alberto Rovellini"
date: "9/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rbgm)
library(sf)
library(viridis)
library(maps)
library(mapdata)
library(data.table)
library(RColorBrewer)
library(lubridate)
library(readxl)
library(ggalluvial)
```

```{r}
select <- dplyr::select
```


This code explores the Fish Ticket data obtained from AKFIN Answers with New --> Analysis --> Fish Tickets.
I seem to have problem downloading the entire data set (possibly too large), so breaking it down in a few 8-yr chunks (TODO: need to do smaller, say 7 years, some days get cut off - all sets have suspiciously same number of rows)

Read data.
```{r}
dat1 <- read.csv('../data/AKFIN/Fish_Tickets/Fish_Tickets_v1_1991-1998.csv', fileEncoding = 'UTF-8-BOM')
dat2 <- read.csv('../data/AKFIN/Fish_Tickets/Fish_Tickets_v1_1999-2006.csv', fileEncoding = 'UTF-8-BOM')
dat3 <- read.csv('../data/AKFIN/Fish_Tickets/Fish_Tickets_v1_2007-2014.csv', fileEncoding = 'UTF-8-BOM')
dat4 <- read.csv('../data/AKFIN/Fish_Tickets/Fish_Tickets_v1_2015-2020.csv', fileEncoding = 'UTF-8-BOM')

dat <- rbind(dat1, dat2, dat3, dat4)
glimpse(dat)
```

Other data and look up tables for gear and targets.
```{r}
# this table is from the eLandings user guide https://elandings.atlassian.net/wiki/spaces/doc/overview
gears <-read_xls('../data/AKFIN/Fish_Tickets/ALLGEAR.XLS', sheet = 1, range = 'A2:I47')
colnames(gears)[1] <- 'Gear.Code'
gears <- gears %>% 
  select(Gear.Code, Description) %>%
  distinct() %>%
  drop_na()


# this table is from the BLEND CAS data, queried as Analysis from AKFIN Answers and only selecting target and gear dimension fields
targets <- read.csv('../data/AKFIN/Fish_Tickets/Target_codes_BLEND.csv', fileEncoding = 'UTF-8-BOM')
targets <- targets %>%
  select(Trip.Target.Code:Trip.Target.Name) %>%
  distinct() %>%
  arrange(Trip.Target.Code)
```

Some data cleaning and general information.
```{r}
# add column for year
datXXX <- dat %>% 
  filter(!is.na(Date.Landed)) %>%
  arrange(Date.Landed) %>%
  distinct() %>%
  mutate(Date.Landed = as.Date(as.character(Date.Landed), '%Y%m%d'),
         Year.Landed = year(Date.Landed))

# join with gear codes to add gear description
datXXX <- datXXX %>%
  left_join(gears, by = 'Gear.Code')

# join with target species 
datXXX <- datXXX %>% 
  left_join(targets, by = c('Groundfish.Weekly.Target'='Trip.Target.Code'))
```

List the ports, the gears, and the targets.
```{r}
datXXX %>% pull(Port.Name) %>% unique() %>% sort()
datXXX %>% pull(Trip.Target.Name) %>% unique() %>% sort()
datXXX %>% pull(Description) %>% unique() %>% sort()
```

For plotting purposes below and general overview, let's extract the dominant ports and species that are reported on the fish ticket.
```{r}
# top ports by landings (annual averages)
top_ports <- datXXX %>%
  group_by(Year.Landed, Port.Name) %>%
  summarise(Landed.t = sum(Pounds..Detail.)*0.454/1000000) %>%
  group_by(Port.Name) %>%
  summarise(Landed.t = mean(Landed.t)) %>%
  arrange(desc(Landed.t)) %>%
  slice(1:25) %>%
  pull(Port.Name)

# top species for average landings per year (regardless of port)
top_species <- datXXX %>%
  group_by(Year.Landed, Species.Common.Name) %>%
  summarise(Landed.t = sum(Pounds..Detail.)*0.454/1000000) %>%
  group_by(Species.Common.Name) %>%
  summarise(Landed.t = mean(Landed.t)) %>%
  arrange(desc(Landed.t)) %>%
  slice(1:25) %>%
  pull(Species.Common.Name)

top_ports
top_species
```

# Ports

What are the main ports in terms of landings? And in terms of revenues? And how does that change over time?
Can this help us with our fleets? Is there anything that can be said about gear / target and port combinations? How do we map ports to the fisheries that we already have? And can we map statistical areas to ports?
```{r}
byport <- datXXX %>%
  filter(Port.Name %in% top_ports) %>%
  group_by(Year.Landed, Port.Name) %>%
  summarise(Landed.t = sum(Pounds..Detail.)*0.454/1000000,
            Revenue = sum(Pounds..Detail.*CFEC.Price.per.Pound)) %>%
  group_by(Port.Name) %>%
  summarise(Landed.t = mean(Landed.t),
            Revenue = mean(Revenue)) %>%
  ungroup() %>%
  arrange(desc(Landed.t)) %>%
  mutate(Port.Name = factor(Port.Name, levels = unique(top_ports))) %>%
  pivot_longer(-Port.Name, names_to = 'Variable', values_to = 'Value')

byport %>%
  ggplot()+
  geom_bar(aes(x = forcats::fct_rev(Port.Name), y = Value), stat = 'identity')+
  coord_flip()+
  theme_bw()+
  facet_wrap(~Variable, scales = 'free_x')+
  labs(x = '', y = 'Port', title = 'Mean annual landings (t) and resulting revenues \n ($) based on CFEC PPP from fish tickets')

```

Kodiak seems to be the port receiving most revenues from the GOA. Some of the catch goes to ports outside the GOA, including Seattle, but it does not seem to be a large proportion. Quite a few blanks and missing/unknown values. Some of the catch goes to floating catcher processors, inshore floating processors, floating domestic motherships.

Temporal changes in total landing per port. Only view the 25 "largest" ports - see above. 
```{r}
byport_temp <- datXXX %>%
  group_by(Year.Landed, Port.Name) %>%
  summarise(Landed.t = sum(Pounds..Detail.)*0.454/1000000)

byport_temp %>%
  filter(Port.Name %in% top_ports) %>%
  ggplot()+
  geom_line(aes(x = Year.Landed, y = Landed.t))+
  theme_bw()+
  facet_wrap(~Port.Name, scales = 'free_y')+
  labs(x = '', y = 'Landings (t)', title = 'Mean annual landings (t)\n')

```

Fluctuations in every port and no clear trend in terms of weight of landings.

Which species are landed the most at each port? 
```{r, fig.width=12, fig.height=8}
# make plots for methods
colourCount <- length(top_species)
getPalette <- colorRampPalette(brewer.pal(12, "Paired"))

byport_species <- datXXX %>%
  filter(Port.Name %in% top_ports, Species.Common.Name %in% top_species) %>%
  group_by(Year.Landed, Port.Name, Species.Common.Name) %>%
  summarise(Landed.t = sum(Pounds..Detail.)*0.454/1000000,
            Revenue = sum(Pounds..Detail.*CFEC.Price.per.Pound)) %>%
  group_by(Port.Name, Species.Common.Name) %>%
  summarise(Landed.t = mean(Landed.t),
            Revenue = mean(Revenue)) %>%
  ungroup() %>%
  arrange(desc(Landed.t)) %>%
  mutate(Port.Name = factor(Port.Name, levels = unique(Port.Name))) %>%
  pivot_longer(-c(Port.Name, Species.Common.Name), names_to = 'Variable', values_to = 'Value')

byport_species %>%
  ggplot()+
  geom_bar(aes(x = Port.Name, y = Value, fill = Species.Common.Name), stat = 'identity', position = position_stack())+
  scale_fill_manual(values = getPalette(colourCount))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 60, hjust = 1, vjust = 1))+
  facet_wrap(~Variable, scales = 'free_y')+
  labs(x = '', y = 'Port', title = 'Mean annual landings (t) and resulting revenues ($) based on CFEC PPP from fish tickets')

```

Same figure but as relative proportions.
```{r, fig.width=12, fig.height=8}
byport_species %>%
  group_by(Port.Name, Variable) %>%
  mutate(Prop = Value/sum(Value)) %>%
  ungroup() %>%
  ggplot()+
  geom_bar(aes(x = Port.Name, y = Prop, fill = Species.Common.Name), stat = 'identity', position = position_stack())+
  scale_fill_manual(values = getPalette(colourCount))+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  facet_wrap(~Variable, scales = 'free_y')+
  labs(x = '', y = 'Port', title = 'Mean annual landings (t) and resulting revenues ($) based on CFEC PPP from fish tickets')
```

# Target-gear and port combinations

Does anything meaningful come up if we combine target, gear, and port of landing?
We use landings instead of revenues as a metric to identify the largest fisheries and ports because in Atlantis we deal mostly with removals (for now).

Careful - target definitions only apply to groundfish.

First - what's with salmon's gear and targets?
```{r}
datXXX %>%
  mutate(Target.Gear = paste(Trip.Target.Name, Description, sep = ' - ')) %>%
  filter(grepl('salmon', Species.Common.Name)) %>%
  group_by(Trip.Target.Group) %>%
  tally() %>%
  arrange(desc(n)) %>%
  mutate(prop = n/sum(n)*100)

datXXX %>%
  mutate(Target.Gear = paste(Trip.Target.Name, Description, sep = ' - ')) %>%
  filter(Trip.Target.Name == '') %>%
  group_by(Species.Common.Name) %>%
  tally() %>%
  arrange(desc(n)) %>%
  mutate(prop = n/sum(n)*100)
```
99% of the landed salmon is labeled with Trip.Target.Name = " ", and 'Trip.Target.Group = "No Retained Groundfish". It is a substantial part of the data, so keep this. Replacing ' ' with salmon would be misleading - I do not understand whether salmon is ever targeted. 

All non-groundfish and state-managed species, including the crabs, are labeled with Trip.Target.Name = " ". We need to tease those apart but there are a lot.

Only consider top 25 ports in terms of landings. Assign each ticket to a decade too, and use decades for summaries (as facets probably).
```{r}
target_gear_combo <- datXXX %>%
  filter(Port.Name %in% top_ports) %>%
  rowwise() %>%
  mutate(Period = ifelse(Year.Landed < 2001, '1991-2000', 
                         ifelse(Year.Landed < 2011, '2001-2010', '2011-2020'))) %>%
  ungroup() %>%
  #filter(Groundfish.Weekly.Target != '') %>% # this removes salmon
  mutate(Trip.Target.Name = na_if(Trip.Target.Name, ''),
         Target.Gear = paste(Trip.Target.Name, Description, sep = ' - ')) %>%
  group_by(Period, Port.Name, Year.Landed, Target.Gear, Species.Common.Name) %>%
  summarise(Landed.t = sum(Pounds..Detail.)*0.454/1000000) %>%
  group_by(Period, Port.Name, Target.Gear, Species.Common.Name) %>%
  summarise(Landed.t = mean(Landed.t)) %>%
  arrange(Period, Port.Name, desc(Landed.t)) %>%
  group_by(Period, Port.Name) %>%
  mutate(Prop.Landings = Landed.t/sum(Landed.t),
         Cumu.Prop = cumsum(Prop.Landings)) %>%
  ungroup() %>%
  filter(Cumu.Prop < 0.95)
```

```{r}
# make some plots as output - one per target-gear combination for readability
these_combos <- target_gear_combo %>% pull(Target.Gear) %>% unique()

for(i in 1: length(these_combos)) {
  
  # write this without slash for the path name when saving the image
  for_path <- gsub('/','--',these_combos[i])
  
  # pull data for this species only
  this_set <- target_gear_combo %>%
    filter(Target.Gear == these_combos[i])
  
  # what species do we have?
  these_species <- this_set %>% pull(Species.Common.Name) %>% unique()
  
  # change colors - they will be different between plots
  colourCount <- length(these_species)
  getPalette <- colorRampPalette(brewer.pal(12, "Paired"))
  
  p <- this_set %>%
    ggplot()+
    geom_bar(aes(x = Port.Name, y = Landed.t, fill = Species.Common.Name), stat = 'identity')+
    scale_fill_manual(values = getPalette(colourCount))+
    theme_bw()+
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
    labs(title = these_combos[i], x = 'Port of landing', y = 'Annual landings (t)')+
    facet_wrap(~Period)
  
  ggsave(paste0('../output/graphics/fleets/from_fish_tickets/by_fishery/',for_path,'.png'),p,width = 12,height = 6)
  
}
```

Do the same but by port instead of by fishery.
```{r}
# make some plots as output - one per target-gear combination for readability
these_ports <- target_gear_combo %>% pull(Port.Name) %>% unique()

for(i in 1: length(these_ports)) {
  
  # write this without slash for the path name when saving the image
  for_path <- gsub('/','--',these_ports[i])
  
  # pull data for this species only
  this_set <- target_gear_combo %>%
    filter(Port.Name == these_ports[i])
  
  # what species do we have?
  these_species <- this_set %>% pull(Species.Common.Name) %>% unique()
  
  # change colors - they will be different between plots
  colourCount <- length(these_species)
  getPalette <- colorRampPalette(brewer.pal(12, "Paired"))
  
  p <- this_set %>%
    ggplot()+
    geom_bar(aes(x = Target.Gear, y = Landed.t, fill = Species.Common.Name), stat = 'identity')+
    scale_fill_manual(values = getPalette(colourCount))+
    theme_bw()+
    theme(axis.text.x = element_text(angle = 60, hjust = 1))+
    labs(title = these_ports[i], x = 'Target-gear combination', y = 'Annual landings (t)')+
    facet_wrap(~Period)
  
  ggsave(paste0('../output/graphics/fleets/from_fish_tickets/by_port/',for_path,'.png'),p,width = 14,height = 8)
  
}
```

# Spatial features of reported fishing

First, we need a map of the statistical areas. For groundfish, there is a downloadable shapefile on the [ADF&G website](https://alaska-department-of-fish-and-game-adfg.hub.arcgis.com/datasets/groundfish-statistical-areas-2001/explore?location=27.354454%2C0.000000%2C2.66). 

```{r}
gf_areas <- st_read('../data/AKFIN/Fish_Tickets/Groundfish_Statistical_Areas_2001/PVG_Statewide_2001_Present_GCS_WGS1984.shp')

gf_areas <- gf_areas %>%
  filter(FMP_AREA_C == 'GOA') %>%
  select(STAT_AREA)

gf_areas %>% ggplot()+geom_sf()+theme_bw()
```

Join this with the ticket data. Aggregate landings by year and stat area (all species).
```{r}
tmp1 <- datXXX %>%
  filter(Groundfish.Weekly.Target != '') %>% # this removes salmon, and we need it here because the ADF&G GF areas only apply to gf
  rowwise() %>%
  mutate(Period = ifelse(Year.Landed < 2001, '1991-2000', 
                         ifelse(Year.Landed < 2011, '2001-2010', '2011-2020'))) %>%
  ungroup() %>%
  mutate(Trip.Target.Name = na_if(Trip.Target.Name, ''),
         Target.Gear = paste(Trip.Target.Name, Description, sep = ' - ')) %>%
  group_by(Period, Statistical.Area, Year.Landed, Target.Gear) %>% #, Species.Common.Name) %>%
  summarise(Landed.t = sum(Pounds..Detail.)*0.454/1000000) %>%
  group_by(Period, Statistical.Area, Target.Gear) %>% #, Species.Common.Name) %>%
  summarise(Landed.t = mean(Landed.t)) %>%
  arrange(Period, Statistical.Area, desc(Landed.t)) %>%
  group_by(Period, Statistical.Area) %>%
  mutate(Prop.Landings = Landed.t/sum(Landed.t),
         Cumu.Prop = cumsum(Prop.Landings)) %>%
  ungroup() %>%
  filter(Cumu.Prop < 0.95)

target_gear_areas <- gf_areas %>%
  full_join(tmp1, by = c('STAT_AREA'='Statistical.Area')) %>%
  filter(!is.na(Period))
```

View in space. Limit this to the target-gear combinations we explored above.
```{r}
# add coastline
obs_bbox <- target_gear_areas %>% st_bbox()

coast <- maps::map("worldHires", c("USA","Canada"), plot = FALSE, fill = TRUE)
coast_sf <- coast %>% st_as_sf() #%>% st_transform(crs = atlantis_crs)

t <- target_gear_areas %>%
  filter(Target.Gear == "NA - Purse seine")

for(i in 1:length(these_combos)){
  
  #print(paste('doing',these_combos[i]))
  
  # write this without slash for the path name when saving the image
  for_path <- gsub('/','--',these_combos[i])
  
  # pull data for this species only
  this_set <- target_gear_areas %>%
    filter(Target.Gear == these_combos[i])
  
  if(nrow(this_set)>0){
    p <- this_set %>%
      ggplot()+
      geom_sf(aes(fill = Landed.t))+
      geom_sf(data = coast_sf)+
      coord_sf(xlim = c(obs_bbox$xmin,obs_bbox$xmax), ylim = c(obs_bbox$ymin,obs_bbox$ymax))+
      scale_fill_viridis()+
      theme_bw()+
      facet_wrap(~Period, ncol = 1)+
      labs(title = these_combos[i])
    
    ggsave(paste0('../output/graphics/fleets/from_fish_tickets/by_area/',for_path,'.png'),p,width = 8,height = 12)
    
  }
  
}
```

TODO: We need salmon statistical areas from ADF&G.

Let's map the ports and the NMFS areas (although those are not useful for salmon).
```{r, fig.width=12, fig.height=6}
#write.csv(data.frame('Port'=top_ports), '../data/AKFIN/Fish_Tickets/ports.csv', row.names = F)
ports <- read.csv('../data/AKFIN/Fish_Tickets/ports_coords.csv')

# NMFS areas
nmfs_areas <- st_read('../data/gf95_nmfs/gf95_nmfs.shp')
nmfs_areas <- nmfs_areas %>% filter(NMFS_AREA %in% c(610,620,630,640,650))

ports <- ports %>% 
  drop_na() %>%
  st_as_sf(coords = c('Lon', 'Lat'), crs = 4326)

port_bbox <- ports %>% st_bbox()


ggplot()+
  geom_sf(data = coast_sf)+
  geom_sf(data = ports, size = 5, color = 'magenta')+
  geom_sf(data = nmfs_areas, color = 'black', fill = NA)+
  ggrepel::geom_label_repel(
    data = ports,
    aes(label = Port, geometry = geometry),
    stat = "sf_coordinates",
    min.segment.length = 0,
    label.size = 0.1,
    segment.colour = 'magenta'
  )+
  theme_bw()+
  coord_sf(xlim = c(port_bbox$xmin,port_bbox$xmax), ylim = c(port_bbox$ymin,port_bbox$ymax))+
  labs(title = 'AK ports with highest mean annual landings (1991-2020) based on Fish Ticket data',
       x = '', y = '')

```

Do the fisheries and their areas map to specific ports? We have a few way to look at this.

Add up landings by NMFS area and by port and see which ports dominate which stat area. Limit this to the top ports.
```{r}
nmfs_to_port <- datXXX %>%
  filter(Port.Name %in% top_ports) %>%
  rowwise() %>%
  mutate(Period = ifelse(Year.Landed < 2001, '1991-2000', 
                         ifelse(Year.Landed < 2011, '2001-2010', '2011-2020'))) %>%
  ungroup() %>%
  # mutate(Trip.Target.Name = na_if(Trip.Target.Name, ''),
  #        Target.Gear = paste(Trip.Target.Name, Description, sep = ' - ')) %>%
  group_by(Period, Reporting.Area.Code, Port.Name, Year.Landed) %>% 
  summarise(Landed.t = sum(Pounds..Detail.)*0.454/1000000) %>%
  group_by(Period, Reporting.Area.Code, Port.Name) %>%
  summarise(Landed.t = mean(Landed.t)) 

# #%>%
#   arrange(Period, Statistical.Area, desc(Landed.t)) %>%
#   group_by(Period, Statistical.Area) %>%
#   mutate(Prop.Landings = Landed.t/sum(Landed.t),
#          Cumu.Prop = cumsum(Prop.Landings)) %>%
#   ungroup() %>%
#   filter(Cumu.Prop < 0.95)

```

View.
```{r, fig.width=12, fig.height=11}
these_periods <- nmfs_to_port %>% pull(Period) %>% unique()

# make plots for methods
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

for(i in 1:length(these_periods)){
  
  p <- nmfs_to_port %>%
    filter(Reporting.Area.Code != 680) %>% #don't know what this is but throw it away
    filter(Period == these_periods[i]) %>%
    ggplot(aes(y = Landed.t, axis1 = Reporting.Area.Code, axis2 = Port.Name))+
    geom_alluvium(aes(fill = factor(Reporting.Area.Code)), width = 1/16)+
    scale_fill_manual(values=cbPalette)+
    geom_stratum(width = 1/16, fill = "white", color = "grey")+
    scale_x_discrete(breaks = NULL)+
    theme_bw()+
    theme(legend.position="none")+
    ggrepel::geom_text_repel(
      aes(label = ifelse(after_stat(x) == 1, as.character(after_stat(stratum)), "")),
      stat = "stratum", size = 4, direction = "y", nudge_x = -3) +
    ggrepel::geom_text_repel(
      aes(label = ifelse(after_stat(x)  == 2, as.character(after_stat(stratum)), "")),
      stat = "stratum", size = 4, direction = "y", nudge_x = 3, force = 4
    )+
    labs(title = paste('Mean annual landings for',these_periods[i],'based on Fish Tickets'),
         y = 'Landings (t)')
  
  ggsave(paste0('../output/graphics/fleets/from_fish_tickets/alluvial/NMFS_to_port_',these_periods[i],'.png'),p,width=11,height=10)
  
}

```

See where the target-gear combinations go to. It is a big list and we need to reduce it for visualisation purposes.
```{r}
targetgear_to_port <- datXXX %>%
  filter(Port.Name %in% top_ports) %>%
  rowwise() %>%
  mutate(Period = ifelse(Year.Landed < 2001, '1991-2000', 
                         ifelse(Year.Landed < 2011, '2001-2010', '2011-2020'))) %>%
  ungroup() %>%
  mutate(Trip.Target.Name = na_if(Trip.Target.Name, ''),
         Target.Gear = paste(Trip.Target.Name, Description, sep = ' - ')) %>%
  group_by(Period, Target.Gear, Port.Name, Year.Landed) %>% 
  summarise(Landed.t = sum(Pounds..Detail.)*0.454/1000000) %>%
  group_by(Period, Target.Gear, Port.Name) %>%
  summarise(Landed.t = mean(Landed.t)) 
```

Drop for viewing purposes and plot.
```{r, fig.width=12, fig.height=13}
these_periods <- targetgear_to_port %>% pull(Period) %>% unique()

top_combos <- targetgear_to_port %>%
  group_by(Period) %>%
  arrange(Period, desc(Landed.t)) %>%
  mutate(Prop = Landed.t/sum(Landed.t),
         Cumu = cumsum(Prop)) %>%
  ungroup() %>%
  filter(Cumu < 0.95) %>%
  pull(Target.Gear) %>%
  unique()

for(i in 1:length(these_periods)){
  
  # make plots for methods
  colourCount <- length(top_combos)
  getPalette <- colorRampPalette(brewer.pal(12, "Paired"))
  
  p <- targetgear_to_port %>%
    filter(Target.Gear %in% top_combos) %>% 
    filter(Period == these_periods[i]) %>%
    ggplot(aes(y = Landed.t, axis1 = Target.Gear, axis2 = Port.Name))+
    geom_alluvium(aes(fill = factor(Target.Gear)), width = 1/16)+
    scale_fill_manual(values = getPalette(colourCount))+
    geom_stratum(width = 1/16, fill = "white", color = "grey")+
    theme_bw()+
    scale_x_discrete(breaks = NULL)+
    ggrepel::geom_text_repel(
      aes(label = ifelse(after_stat(x) == 1, as.character(after_stat(stratum)), "")),
      stat = "stratum", size = 4, direction = "y", nudge_x = -3, force = 4) +
    ggrepel::geom_text_repel(
      aes(label = ifelse(after_stat(x)  == 2, as.character(after_stat(stratum)), "")),
      stat = "stratum", size = 4, direction = "y", nudge_x = 3, force = 4
    )+
    theme(legend.position="none")+
    labs(title = paste('Mean annual landings for',these_periods[i],'based on Fish Tickets'),
         y = 'Landings (t)')
  
  ggsave(paste0('../output/graphics/fleets/from_fish_tickets/alluvial/targetgear_to_port_',these_periods[i],'.png'),p,width=12,height=10)
  
}

```

Let's zoom into one target-gear combination at a time, if need be, and have a look at the species composition of the landings.
```{r}
targetgear_to_port_species <- datXXX %>%
  filter(Port.Name %in% top_ports, Species.Common.Name %in% top_species) %>%
  rowwise() %>%
  mutate(Period = ifelse(Year.Landed < 2001, '1991-2000', 
                         ifelse(Year.Landed < 2011, '2001-2010', '2011-2020'))) %>%
  ungroup() %>%
  mutate(Trip.Target.Name = na_if(Trip.Target.Name, ''),
         Target.Gear = paste(Trip.Target.Name, Description, sep = ' - ')) %>%
  group_by(Period, Target.Gear, Port.Name, Year.Landed, Species.Common.Name) %>% 
  summarise(Landed.t = sum(Pounds..Detail.)*0.454/1000000) %>%
  group_by(Period, Target.Gear, Port.Name, Species.Common.Name) %>%
  summarise(Landed.t = mean(Landed.t)) 
```

Drop for viewing purposes and plot.
```{r, fig.width=12, fig.height=13}
these_periods <- targetgear_to_port_species %>% pull(Period) %>% unique()
these_combos <-  targetgear_to_port_species %>% pull(Target.Gear) %>% unique()

for(i in 1:length(these_periods)){
  
  for(j in 1:length(these_combos)){
    
    this_set <- targetgear_to_port_species %>%
      filter(Period == these_periods[i], Target.Gear == these_combos[j])
    
    # write this without slash for the path name when saving the image
    for_path <- gsub('/','--',these_combos[j])
    
    if(nrow(this_set)>0){
      # make plots for methods
    colourCount <- length(this_set %>% pull(Species.Common.Name) %>% unique())
    getPalette <- colorRampPalette(brewer.pal(12, "Paired"))
    
    p <-  this_set %>%
      ggplot(aes(y = Landed.t, axis1 = Species.Common.Name, axis2 = Port.Name))+
      geom_alluvium(aes(fill = factor(Species.Common.Name)), width = 1/16)+
      scale_fill_manual(values = getPalette(colourCount))+
      geom_stratum(width = 1/16, fill = "white", color = "grey")+
      theme_bw()+
      scale_x_discrete(breaks = NULL)+
      ggrepel::geom_text_repel(
        aes(label = ifelse(after_stat(x) == 1, as.character(after_stat(stratum)), "")),
        stat = "stratum", size = 4, direction = "y", nudge_x = -3, force = 4) +
      ggrepel::geom_text_repel(
        aes(label = ifelse(after_stat(x)  == 2, as.character(after_stat(stratum)), "")),
        stat = "stratum", size = 4, direction = "y", nudge_x = 3, force = 4
      )+
      #theme(legend.position="none")+
      labs(title = paste('Mean annual landings for',these_periods[i],these_combos[j],'\nbased on Fish Tickets'),
           y = 'Landings (t)')
    
    p
    
    ggsave(paste0('../output/graphics/fleets/from_fish_tickets/alluvial/targetgear_to_port_species/',
                  these_periods[i],for_path,'.png'),p,width=12,height=10)
    }
    
  }
  
}
```
